\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2021-12-03 main.Rmd
more tweaks of text on examples

**Our new data structure for spatio-temporal data**  
This paper describes the implementation of a new spatio-temporal data structure: `cubble`. `cubble` implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension. With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to `cubble`. The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link]. \newline

@@

\newpage
# The cubble package

@@

\begin{center}
\begin{tabular}{ | m{5em} | m{15cm}| } 
basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
grouping &  \textbf{group\_by}, \textbf{ungroup}\\
slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
\end{tabular}
\end{center}

`cubble` is also compatible with `tsibble` in the sense that the original list-column can be a `tbl_ts` object. Duplicates and gaps should be first checked before structuring the data into a cubble. If the input data is a `tsibble` object, the long form cubble is also a `tsibble` where users can directly apply time series operations.

@@

@@

The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks. Then you can use this output to plot the location of match or to look at the series: 
```{r echo = FALSE, out.width="100%", fig.cap="..."}
p1 <- plot_map(vic_map) +
  geom_point(
    data = res,
    aes(x = long, y = lat, color = type)
  ) +
  ggrepel::geom_label_repel(
    data = res %>% filter(type == "river"),
    aes(x = long, y = lat, label = .group)
  ) +
  scale_color_brewer(palette = "Dark2") +
  ggtitle("Victoria") +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )

@@

@@

```{r eval = FALSE}
set.seed(123)
dt <- weatherdata::climate_full %>%
  slice_sample(n = 50) %>% 
  mutate(name = str_replace(name, "\\s*\\([^\\)]+\\)", ""))

tmax_data <- dt %>%
  stretch() %>%
  mutate(ym = tsibble::yearmonth(date)) %>%
  group_by(ym) %>%
  summarise(
    tmax = mean(tmax, na.rm = TRUE),
    tmin = mean(tmin, na.rm = TRUE),
    prcp = sum(prcp, na.rm = TRUE)
  ) %>%
  ungroup(ym) %>%
  mutate(
    year = as.factor(year(ym)),
    month = as.factor(month(ym)),
    prcp = sqrt(prcp + 0.001)
  ) %>%
  tamp() %>%
  mutate(median_tmax = median(ts$tmax, na.rm = TRUE)) %>%
  stretch() %>%
  migrate(median_tmax, long, lat, name) %>%
  highlight_key(~id)
```

```{r eval = FALSE, echo = FALSE}
p1 <- tmax_data %>%
  ggplot(aes(
    x = month, y = tmax, group = interaction(year, id),
    color = median_tmax, label = name
  )) +
  geom_line() +
  scale_x_discrete(breaks = seq(1, 12, 3)) + 
  facet_wrap(vars(fct_reorder(name, -median_tmax))) +
  theme(legend.position = "none")

state_map <- rmapshaper::ms_simplify(ozmaps::abs_ste, keep = 2e-3)
p2 <- tmax_data %>%
  ggplot() +
  geom_sf(data = state_map, aes(geometry = geometry)) +
  geom_point(aes(x = long, y = lat, color = median_tmax, label = name)) +
  theme_void()

bscols(
  ggplotly(p1, width = 800, height = 800, tooltip = "label") %>%
    highlight(on = "plotly_selected", off = "plotly_deselect", color = "red"),
  ggplotly(p2, width = 800, height = 800, tooltip = "label") %>%
    highlight(on = "plotly_selected", off = "plotly_deselect", color = "red")
)
```

>>>>

**Our new data structure for spatio-temporal data**\
This paper describes the implementation of a new spatio-temporal data structure: `cubble`. `cubble` implements a relational data structure that uses two forms to manage the switch between spatial and temporal dimension. With this structure, users can manipulate the spatial or temporal dimension separately, while leaves the linking of two dimensions to `cubble`. The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link]. \newline

@@

# The cubble package

@@

```{=tex}
\begin{center}
\begin{tabular}{ | m{5em} | m{15cm}| } 
basics & \textbf{mutate}, \textbf{filter}, \textbf{summarise}, \textbf{select}, \textbf{arrange}, \textbf{rename}, \textbf{left\_join} \\
grouping &  \textbf{group\_by}, \textbf{ungroup}\\
slice family & \textbf{slice\_head}, \textbf{slice\_tail}, \textbf{slice\_sample}, \textbf{slice\_min} and \textbf{slice\_max} \\
\end{tabular}
\end{center}
```
`cubble` is also compatible with `tsibble` in the sense that the original list-column can be a `tbl_ts` object. Duplicates and gaps should be first checked before structuring the data into a cubble. If the input data is a `tsibble` object, the long form cubble is also a `tsibble` where users can directly apply time series operations.

@@

Interactive graphics give users the chance to manipulate the graphes for additional information. This is especially useful for spatio-temporal data since it would be interesting to explore the correlation between nearby sites on the map or compare between clusters. Underlying each interactive graphic system is a pipeline workflow that communicates the user action down the workflow to [...]. Various such workflow has been proposed in the 90s literature, including [...]. Current general purpose R implementation on iterative graphics includes: plotly [@plotly] for [...]; ggiraph [@ggiraph] for [...]; crosstalk[@crosstalk] for [...]. Specifically for spatio-temporal data, there is tmap [@map], leftlet [@leaflet], mapview [@mapview].

some common interactive actions include zoom and panning, tooltip/ highlight, and more advanced cross-linking.

@@

The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks. Then you can use this output to plot the location of match or to look at the series:

```{r echo = FALSE, out.width="100%", fig.cap="..."}
p1 <- plot_map(vic_map) +
  geom_point(
    data = res,
    aes(x = long, y = lat, color = type)
  ) +
  ggrepel::geom_label_repel(
    data = res %>% filter(type == "river"),
    aes(x = long, y = lat, label = .group)
  ) +
  scale_color_brewer(palette = "Dark2") +
  ggtitle("Victoria") +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )

@@

## Spatio-temporal data with NetCDF

```{r}
nc <- ncdf4::nc_open(here::here("data/adaptor.mars.internal-1638344933.932894-4194-11-a5d35ac3-b36d-4982-afa5-a5dd193a6718.nc"))
temp <- as_cubble(nc, vars = "d2m")
vic_map <- ozmaps::abs_ste %>% filter(NAME == "Victoria")

res <- temp %>% 
  stretch(ts) %>% 
  mutate(d2m = d2m - 273.15) %>% 
  filter(lubridate::hour(time) == 12) %>% 
  group_by(id) %>% 
  summarise(d2m = mean(d2m, na.rm= TRUE)) %>%
  migrate(long, lat)

vic_string <- vic_map %>% sf::st_cast("MULTILINESTRING")

res %>% 
  ggplot() +
  geom_tile(aes(x = long, y = lat, fill = d2m)) +
  geom_sf(data = vic_string , alpha = 0.5, fill = "transparent") +
  theme_bw() + 
  colorspace::scale_fill_continuous_sequential("Mint")

```

@@

```{r echo = FALSE, out.width="100%", out.height="30%"}
knitr::include_graphics(here::here("figures/linking.png"))
```

### 2021-12-03 main.tex
more tweaks of text on examples

@@

\begin{CodeChunk}
\begin{CodeInput}
R> set.seed(123)
R> dt <- weatherdata::climate_full %>%
+   slice_sample(n = 50) %>% 
+   mutate(name = str_replace(name, "\\s*\\([^\\)]+\\)", ""))
R>  
R> tmax_data <- dt %>%
+   stretch() %>%
+   mutate(ym = tsibble::yearmonth(date)) %>%
+   group_by(ym) %>%
+   summarise(
+     tmax = mean(tmax, na.rm = TRUE),
+     tmin = mean(tmin, na.rm = TRUE),
+     prcp = sum(prcp, na.rm = TRUE)
+   ) %>%
+   ungroup(ym) %>%
+   mutate(
+     year = as.factor(year(ym)),
+     month = as.factor(month(ym)),
+     prcp = sqrt(prcp + 0.001)
+   ) %>%
+   tamp() %>%
+   mutate(median_tmax = median(ts$tmax, na.rm = TRUE)) %>%
+   stretch() %>%
+   migrate(median_tmax, long, lat, name) %>%
+   highlight_key(~id)
\end{CodeInput}
\end{CodeChunk}

>>>>

\hypertarget{spatio-temporal-data-with-netcdf}{%
\subsection{Spatio-temporal data with
NetCDF}\label{spatio-temporal-data-with-netcdf}}

\begin{CodeChunk}
\begin{CodeInput}
R> nc <- ncdf4::nc_open(here::here("data/adaptor.mars.internal-1638344933.932894-4194-11-a5d35ac3-b36d-4982-afa5-a5dd193a6718.nc"))
R> temp <- as_cubble(nc, vars = "d2m")
R> vic_map <- ozmaps::abs_ste %>% filter(NAME == "Victoria")
R> 
R> res <- temp %>% 
+   stretch(ts) %>% 
+   mutate(d2m = d2m - 273.15) %>% 
+   filter(lubridate::hour(time) == 12) %>% 
+   group_by(id) %>% 
+   summarise(d2m = mean(d2m, na.rm= TRUE)) %>%
+   migrate(long, lat)
R> 
R> vic_string <- vic_map %>% sf::st_cast("MULTILINESTRING") 
R> 
R> res %>% 
+   ggplot() +
+   geom_tile(aes(x = long, y = lat, fill = d2m)) +
+   geom_sf(data = vic_string , alpha = 0.5, fill = "transparent") +
+   theme_bw() + 
+   colorspace::scale_fill_continuous_sequential("Mint")
\end{CodeInput}

\begin{center}\includegraphics{figures/unnamed-chunk-23-1} \end{center}

@@

\begin{center}\includegraphics[width=1\linewidth,height=0.3\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/linking} \end{center}
